---
title: "scorecard_analysis"
author: "Sumitra, Jamie, and Naeem"
date: "October 8, 2019"
output: html_document
---

##Setup
### options
Set up global options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```


### libraries
Load in needed libraries 
```{r libraries}
# Run these if you don't have these libraries yet
# install.packages("tidyverse") 
# install.packages("data.table")
library(tidyverse) # For data processing and visualization
library(data.table) # For creating tables

spot_color = "#9ecae1" 
text_color = "#525252"
```

### File management
Create variables for directories
```{r file_management}
home.dir <- getwd()
data.dir <- '../CollegeScorecard_Raw_Data'
output.dir <- '../outputs'
```

## Preprocessing
### import
This chunk will eventually read in data for all years. Will get completed after we've settled on subsets. 
__TO DO__
- coerce needed columns to correct data type
```{r preprocessing}
import_all <- function() {
    list.files(pattern = "*.csv") %>% 
    map_df(~read_csv(., na = c("NULL")))
}
```

### Test Import for small dataset
```{r}

scorecard_14 <- read_csv(file.path(data.dir, 'MERGED2014_15_PP.csv'), na = c("NULL", "PrivacySuppressed")) 
scorecard_15 <- read_csv(file.path(data.dir, 'MERGED2015_16_PP.csv'), na = c("NULL", "PrivacySuppressed")) 
scorecard_16 <- read_csv(file.path(home.dir, data.dir, "MERGED2016_17_PP.csv"), na = c("NULL", "PrivacySuppressed"))
scorecard_17 <- read_csv(file.path(home.dir, data.dir, "MERGED2017_18_PP.csv"), na = c("NULL", "PrivacySuppressed"))
```

### Subset and merge
```{r}

multmerge = function(mypath){
  #years = c("96_97","97_98", "98_99", "99_00", "00_01", "01_02", "02_03", "03_04", "04_05", "05_06", "06_07", "07_08", "08_09", "09_10", "10_11", "11_12", "12_13", "13_14", "14_15", "15_16", "16_17", "17_18")
  filenames=list.files(path=mypath, pattern = "*.csv") %>% 
  map_df(~read_csv(., col_types = cols_only(INSTNM = col_character(), HCM2 = col_integer(), PREDDEG = col_integer(), HIGHDEG = col_integer(), REGION = col_integer(), PFTFAC = col_number(), AVGFACSAL = col_number(), PCIP05 = col_number(), PCIP16 = col_number(), PCIP27 = col_number()), na = c("", "NA", 0.0000)), .id = 'File')
}



setwd(data.dir)
mergeData <- multmerge(data.dir)
setwd(home.dir)
```


### Quality Measurements
```{r}
# What is the proportion of nulls in the MD_EARN_WNE_P10 column?
length(scorecard_17$MD_EARN_WNE_P10[scorecard_17$MD_EARN_WNE_P10 == 'NULL']) /nrow(scorecard_17)

# Since the proportion of nulls is 1, all of the rows in the column are NULL.
```

### Proportion of NA and duplicates in columns
Below is an example of a function to generate data quality measures on every column. Let's do this after subset. 
```{r}
na_dupe_prop <- function(data, tag) {
  
    reportList <- data_frame() # initialize empty data frame
    
    ##### Determine the proportion of NAs in each column
    # Count all NAs in all columns
    countNA <- apply(data, 2, function(x) sum(is.na(x)))
    # Calculate the proportion of all column values that are NA
    propNA <- round(countNA / nrow(data), 2)

    ##### Determine the number of rows that are duplicated
    # Count all duplicate values in each column
    countDuples <- apply(data, 2, function(x) sum(duplicated(x)))
    # Calculate the proportion of all column values that are duplicates
    propDuples <- round(countDuples / nrow(data), 2)

    reportList <- data.frame(cbind(names(propNA), propNA, propDuples))

    write_csv(reportList, file.path(output.dir, paste0("na_dupe_prop", tag, ".csv")))
}

na_dupe_prop(scorecard_17, "17_18")
```

## Exploration (Playground)

In this exploration section, we can use the most recent scorecard data to look into what parts of the data are interesting before we look over all years.
### How much do schools really spend on their students?
```{r}
selective_schools <- scorecard_17 %>%
    select(INSTNM, TUITIONFEE_IN, DEBT_MDN, UGDS_BLACK, UGDS_WHITE, INEXPFTE, ADM_RATE, AVGFACSAL) %>%
    filter(ADM_RATE < .25)

unselective_schools <- scorecard_17 %>%
    select(INSTNM, TUITIONFEE_IN, DEBT_MDN, UGDS_BLACK, UGDS_WHITE, INEXPFTE, ADM_RATE, AVGFACSAL
) %>%
    filter(ADM_RATE < .95)


# Only at the most expensive selective schools does per-student spending really go up with tuition 
plot(selective_schools$TUITIONFEE_IN, selective_schools$INEXPFTE)

# It's very similar at the less selective schools
plot(unselective_schools$TUITIONFEE_IN, unselective_schools$INEXPFTE)
```

### CC Debt vs. Future Earnings by Control Type
```{r}
debt_v_income <- scorecard_14 %>%
    filter(PREDDEG == 2) %>%
    select(DEBT_MDN, MD_EARN_WNE_P8, CONTROL) %>%
    mutate(CONTROL = factor(CONTROL, labels = c("Public", "Private nonprofit", "Private for-profit"))) %>%
    drop_na()

p <- ggplot(debt_v_income, aes(DEBT_MDN, MD_EARN_WNE_P8)) +
    geom_point(colour=spot_color, size = 2) +
    facet_wrap(~CONTROL, dir = "v") +
    theme(plot.title = element_text(color=text_color, face="bold", size=12, hjust=.5, lineheight=.65)) +
    geom_smooth(method="loess", se=FALSE) +
    scale_y_log10(labels = scales::comma) +
    labs(title="Relationship Between Size of Loan and \n Earnings 8 Years After Graduation \n By Type of School \n (Associate degrees only)",
         x="Size of Original Loan", y="Earnings 8 Years After Graduation") 
    
p
cor(debt_v_income$DEBT_MDN, debt_v_income$MD_EARN_WNE_P8)

```
### Debt vs. Accrediting Agency
```{r fig.width=8, fig.height=10, echo=FALSE}
debt_v_accred <- scorecard_14 %>%
    select(-ACCREDAGENCY) %>%
    left_join(select(scorecard_17, ACCREDAGENCY, UNITID), by = "UNITID") %>%
    select(INSTNM, CONTROL, DEBT_MDN, ACCREDAGENCY) %>%
    mutate(ACCREDAGENCY = as.factor(ACCREDAGENCY)) %>%
    drop_na()

debt_v_accred_plot <- ggplot(debt_v_accred, aes(x=ACCREDAGENCY, y=DEBT_MDN)) + 
        geom_boxplot() +
        labs(title="Debt by Accrediting Agency",
             x="Agency", y = "Median Debt at Graduation") +
        coord_flip() +
        theme(plot.title = element_text(family = "Helvetica", color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1))
debt_v_accred_plot
```

### Loan debt over regions
```{r}
debt_v_region <- scorecard_14 %>% 
  select(-REGION) %>% 
  left_join(select(scorecard_16, REGION, UNITID), by = "UNITID") %>% 
  select(INSTNM, CONTROL, DEBT_MDN, REGION) %>% 
  mutate(REGION = as.factor(REGION)) %>% 
  drop_na()


levels(debt_v_region$REGION) <- c("U.S. Service Schools",
                                  "New England (CT, ME, MA, NH, RI, VT)",
                                  "Mid East (DE, DC, MD, NJ, NY, PA)",
                                  "Great Lakes (IL, IN, MI, OH, WI)",
                                  "Plains (IA, KS, MN, MO, NE, ND, SD)",
                                  "Southeast (AL, AR, FL, GA, KY, LA, MS, NC, SC, TN, VA, WV)",
                                  "Southwest (AZ, NM, OK, TX)",
                                  "Rocky Mountains (CO, ID, MT, UT, WY)",
                                  "Far West (AK, CA, HI, NV, OR, WA)",
                                  "Outlying Areas (AS, FM, GU, MH, MP, PR, PW, VI")


debt_v_region_plot <- ggplot(debt_v_region, aes(x = REGION, y = DEBT_MDN)) +
  geom_boxplot() +
  labs(title="Debt by Region", x="Region", y = "Median Debt at Graduation") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1))

debt_v_region_plot  
```

### 
```{r}
install.packages(c("GGally", "hexbin", "repr", "psych"))

scorecard.df <- scorecard_17 %>% 
  select(INSTNM,CITY,STABBR,ZIP,ACCREDAGENCY,PREDDEG,HIGHDEG,HCM2,CONTROL,REGION,RELAFFIL,COSTT4_A,COSTT4_P,TUITIONFEE_IN,TUITIONFEE_OUT,TUITIONFEE_PROG,OPEID,OPEID6,MAIN,NUMBRANCH,ST_FIPS,CURROPER,ICLEVEL,OPEFLAG) 

#scorecard.mapdf <- scorecard_17 %>% 
  #select(INSTNM,CITY,STABBR,ZIP,ACCREDAGENCY,PREDDEG,HIGHDEG,HCM2,CONTROL,REGION,RELAFFIL,COSTT4_A,COSTT4_P,TUITIONFEE_IN,TUITIONFEE_OUT,TUITIONFEE_PROG,OPEID,OPEID6,MAIN,NUMBRANCH,ST_FIPS,CURROPER,ICLEVEL,OPEFLAG,LATITUDE,LONGITUDE) %>% filter(HCM2==1)

scorecard.df$school_type[scorecard.df$CONTROL == 1] = 'Public'
scorecard.df$school_type[scorecard.df$CONTROL == 2] = 'Private nonprofit'
scorecard.df$school_type[scorecard.df$CONTROL == 3] = 'Private forprofit'
  
scorecard.df$hcm2_status[scorecard.df$HCM2 == 1] = 'Watch_Flag'
scorecard.df$hcm2_status[scorecard.df$HCM2 == 0] = 'No_Flag'

col_nam <- c('PREDDEG','HIGHDEG','ICLEVEL')
            
ggpairs(scorecard.df,
    columns = col_nam,
    aes(color = school_type, alpha = 0.4),
    lower = list(continuous = 'points'),
    upper = list(continuous = 'cor'))

#Tried to map scorecard.mapdf, but not working for now.The map looks weird. 
# install.packages("leaflet")
# library('leaflet')
# leaflet() %>%  addTiles() %>% addMarkers(lng=~LONGITUDE, lat=~LONGITUDE, popup=~sprintf('%s<br/>%s', ZIP, STABBR,CITY), data= scorecard.mapdf)
```


### Tuition vs. Debt
```{r}
TUITIONFEE_IN <- scorecard_16$TUITIONFEE_IN
DEBT_MDN <- scorecard_16$DEBT_MDN
ggplot(scorecard_16, aes(x = TUITIONFEE_IN, y = DEBT_MDN, color= factor(scorecard_16$CONTROL))) + geom_point()
```


### Is it harder for POC to be successful in the South?
```{r}
black_completion_rate <- scorecard_14 %>% 
  select(-REGION) %>% 
  left_join(select(scorecard_16, REGION, UNITID), by = "UNITID") %>% 
  select(INSTNM, 
         CONTROL,
         C150_4_WHITE,
         C150_4_BLACK, 
         C150_4_HISP, 
         C150_4_ASIAN,
         C150_4_AIAN,
         C150_4_NHPI, 
         REGION) %>% 
  mutate(REGION = as.factor(REGION)) %>% 
  drop_na()


levels(black_completion_rate$REGION) <- c("U.S. Service Schools",
                                  "New England (CT, ME, MA, NH, RI, VT)",
                                  "Mid East (DE, DC, MD, NJ, NY, PA)",
                                  "Great Lakes (IL, IN, MI, OH, WI)",
                                  "Plains (IA, KS, MN, MO, NE, ND, SD)",
                                  "Southeast (AL, AR, FL, GA, KY, LA, MS, NC, SC, TN, VA, WV)",
                                  "Southwest (AZ, NM, OK, TX)",
                                  "Rocky Mountains (CO, ID, MT, UT, WY)",
                                  "Far West (AK, CA, HI, NV, OR, WA)",
                                  "Outlying Areas (AS, FM, GU, MH, MP, PR, PW, VI")


black_completion_rate_plot <- ggplot(black_completion_rate, aes(x = REGION, y = C150_4_BLACK)) +
  #geom_violin(aes(fill = REGION)) +
  geom_boxplot(aes(fill = REGION)) +
  labs(title="Black Grad Rate vs Region", x="Region", y = "Grad Rate") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

black_completion_rate_plot  

hispanic_completion_rate_plot <- ggplot(black_completion_rate, aes(x = REGION, y = C150_4_HISP)) +
  #geom_violin(aes(fill = REGION)) +
  geom_boxplot(aes(fill = REGION)) +
  labs(title="Hispanic Grad Rate vs Region", x="Region", y = "Grad Rate") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

hispanic_completion_rate_plot  

asian_completion_rate_plot <- ggplot(black_completion_rate, aes(x = REGION, y = C150_4_ASIAN)) +
  #geom_violin(aes(fill = REGION)) +
  geom_boxplot(aes(fill = REGION)) +
  labs(title="Asian Grad Rate vs Region", x="Region", y = "Grad Rate") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

asian_completion_rate_plot  

white_completion_rate_plot <- ggplot(black_completion_rate, aes(x = REGION, y = C150_4_WHITE)) +
  #geom_violin(aes(fill = REGION)) +
  geom_boxplot(aes(fill = REGION)) +
  labs(title="White Grad Rate vs Region", x="Region", y = "Grad Rate") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

white_completion_rate_plot  
```


### How have things changed over time?
```{r}
mergedSummary <- mergeData %>% 
  group_by(File) %>% 
  summarize(mean_fac_sal = mean(AVGFACSAL, na.rm = TRUE), mean_ethnic_cultural_gender = mean(PCIP05, na.rm = TRUE), mean_math = mean(PCIP27, na.rm = TRUE), mean_language = mean(PCIP16, na.rm=TRUE))



```






